---
title: "Logistic Regression"
author: "Walid Farah"
date: "29/03/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# 1.Loading Different Required Libraries.

```{r warning=FALSE, include=FALSE}
library(leaps)
library(devtools)
library(VIM)
library(pscl)
library(ResourceSelection)
require(HH)
require(pscl)
require(ggplot2)
require(gridExtra)
require(scales)
library(car)


```

# 2.Loading, Preparing and Cleaning the Data.

```{r echo=FALSE, warning=FALSE}
Loans = read.csv("data.csv")
Loans[Loans == "NA"] = NA
sum(is.na(Loans))#sum of the na value
colMeans(is.na(Loans))
rowMeans(is.na(Loans))
Loans = na.omit(Loans)#removing or dropping the NA as it's less than 5% of the Data
Loans[Loans == "NA"] = NA
sum(is.na(Loans))

```

# 3.Exploring and Transforming the Data

```{r echo=FALSE, warning=FALSE}
summary(Loans)
Loans$state=NULL #dropping the State Variable
Loans$loanID = NULL # dropping the Id Variable
Loans$employment = NULL # dropping employemnt variable
summary(Loans)
Loans$status = as.factor(Loans$status)
levels(Loans$status) = list("Good"=c("Fully Paid"),"Bad"=c('Charged Off',"In Grace Period","Late (16-30 days)","Late (31-120 days)"))
Loans$verified = as.factor(Loans$verified)
levels(Loans$verified) = list("Verified"=c("Verified","Source Verified"),"Not Verified"=c("Not Verified"))
sum(is.na(Loans))
Loans = na.omit(Loans)
table(Loans$status)
Loans$status = as.character(Loans$status)
Loans$length = as.character(Loans$length)
Loans$reason = as.character(Loans$reason)
Loans$amount = as.numeric(Loans$amount)
Loans$rate =  as.numeric(Loans$rate)
Loans$term = as.character.Date(Loans$term)
Loans$grade = as.character(Loans$grade)
Loans$length = as.character.Date(Loans$length)
Loans$home = as.character(Loans$home)
Loans$payment =  as.numeric(Loans$payment)
Loans$income =  as.numeric(Loans$income)
Loans$debtIncRat =  as.numeric(Loans$debtIncRat)
Loans$delinq2yr =  as.numeric(Loans$delinq2yr)
Loans$inq6mth =  as.numeric(Loans$inq6mth)
Loans$openAcc =  as.numeric(Loans$openAcc)
Loans$pubRec =  as.numeric(Loans$pubRec)
Loans$revolRatio  =  as.numeric(Loans$revolRatio)
Loans$totalAcc  =  as.numeric(Loans$totalAcc)
Loans$totalPaid  =  as.numeric(Loans$totalPaid)
Loans$totalBal  =  as.numeric(Loans$totalBal)
Loans$totalRevLim  =  as.numeric(Loans$totalRevLim)
Loans$accOpen24 	 =  as.numeric(Loans$accOpen24)
Loans$avgBal  =  as.numeric(Loans$avgBal)
Loans$bcOpen  =  as.numeric(Loans$bcOpen )
Loans$bcRatio =  as.numeric(Loans$bcRatio)
Loans$totalLim  =  as.numeric(Loans$totalLim)
Loans$totalRevBal  =  as.numeric(Loans$totalRevBal)
Loans$totalBcLim  =  as.numeric(Loans$totalBcLim)
Loans$totalIlLim  =  as.numeric(Loans$totalIlLim )
LoansInt = Loans
LoansInt$verified = NULL
LoansInt$term = NULL
LoansInt$status = NULL
LoansInt$grade = NULL
LoansInt$length = NULL
LoansInt$home = NULL
LoansInt$reason = NULL
LoansInt
print(paste("The Profit made by the Bank is :",as.integer(sum(Loans$totalPaid-Loans$amount)),"$"))

```

```{r echo=FALSE, warning=FALSE}
Cormat = cor(LoansInt,use = 'everything',method = "pearson")
Cormat
corrplot::corrplot(Cormat, method="circle", type = "upper",tl.col="black", title="Correlation",mar=c(1,0,1,0))
boxplot(Loans$inq6mth ~ Loans$status, data = Loans,xlab ="Status",ylab="Inquiry in the Last 6months" , col = "lightgray")

ggplot(Loans) +
  aes(x = amount, y = totalPaid, colour = status) +
  geom_point(shape = "circle", size = 1.5) +
  scale_color_hue(direction = 1) +
  labs(x = "Loan Amount", y = "Total Paid", color = "Status") +
  theme_minimal()

ggplot(Loans) +
  aes(x = status, fill = grade) +
  geom_bar() +
  scale_fill_manual(
    values = c(A = "#F8766D",
    B = "#BD9A00",
    C = "#31B425",
    D = "#00C19F",
    E = "#20AFEC",
    F = "#B280FC",
    G = "#FF61C3")
  ) +
  theme_gray()


```

# 4.The Preparing Train and Test dataset for Model

```{r echo=TRUE, warning=FALSE}
samplingsize = floor(0.80 * nrow(Loans)) #to know what is the 80% of my Data
set.seed(123)
samplingsize
traindata = sample(seq_len(nrow(Loans)),size = samplingsize)
TrainLoans = Loans[traindata,] # saving 80% of the data as training data
TrainLoans
TestLoans = Loans[-traindata,]#Saving 20% of the Data in Test dataset
TestLoans
TrainLoans$status = as.factor(TrainLoans$status)#convert to factor variable
```

# 5. LGM "Logistic Regression Model using GLM

```{r echo=TRUE, warning=FALSE}
LrgModel = glm(status ~ .,data = TrainLoans, family = 'binomial')
summary(LrgModel)
prLrg = predict(LrgModel, newdata = TestLoans, type = "response") # to do prediction
trs = 0.5#threshold of 0.5
prediction = cut(prLrg, breaks=c(-Inf,trs, Inf), labels=c("Bad", "Good"), header = TRUE)
PrEdCAL= table(prediction,TestLoans$status)
addmargins(PrEdCAL)

PLevel = round(sum(diag(PrEdCAL)) / sum(PrEdCAL)*100 ,2) # Percentage correctly predicted
bad = round(diag(PrEdCAL)[1]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate correctly predicted bad loans
good = round((PrEdCAL)[4]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate correctly predicted good loans
incbad = round((PrEdCAL)[3]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate incorrectly predicted bad loans
incgood = round((PrEdCAL)[2]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate incorrectly predicted good loans
print(paste0('Correctly predicted loans = ', PLevel,'%'))
print(paste0('Correctly predicted as Bad loans ', bad ,'% ', 'and as Good loans is ',good,'%'))
print(paste0('Incorrectly predicted as bad loans ', incbad,'% ', 'and as good loans is ',incgood,'%'))
```

# 6.Balancing the Data as 50/50 Good/Bad

```{r echo=FALSE, warning=FALSE}
table(TrainLoans$status) # 6681 Bad and 21376 good Loans
Badt = TrainLoans[TrainLoans$status=='Bad',]
Radd = sample(6681, 14695, replace = TRUE)
Balanced = rbind(TrainLoans, Badt[Radd,] )
TrainLoans = Balanced
table(TrainLoans$status)
```

# 7.LGM "Logistic Regression Model in GLM with balanced Data

```{r echo=TRUE, warning=FALSE}

set.seed(123)
LrgModel = glm(status ~ .,data = TrainLoans, family = 'binomial')
summary(LrgModel)
prLrg = predict(LrgModel, newdata = TestLoans, type = "response") # to do prediction
trs = 0.5#threshold of 0.5
prediction = cut(prLrg, breaks=c(-Inf,trs, Inf), labels=c("Bad", "Good"), header = TRUE)
PrEdCAL= table(prediction,TestLoans$status)
addmargins(PrEdCAL)

PLevel = round(sum(diag(PrEdCAL)) / sum(PrEdCAL)*100 ,2) # Percentage correctly predicted
bad = round(diag(PrEdCAL)[1]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate correctly predicted bad loans
good = round((PrEdCAL)[4]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate correctly predicted good loans
incbad = round((PrEdCAL)[3]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate incorrectly predicted bad loans
incgood = round((PrEdCAL)[2]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate incorrectly predicted good loans

print(paste0('Correctly predicted loans = ', PLevel,'%'))
print(paste0('Correctly predicted as Bad loans ', bad ,'% ', 'and as Good loans is ',good,'%'))
print(paste0('Incorrectly predicted as bad loans ', incbad,'% ', 'and as good loans is ',incgood,'%'))
```

# 8.Using Step wise regression Model AIC for Best fit Line.

```{r echo=TRUE, warning=FALSE}
vif(LrgModel)
Fmodel = glm(status ~.,data = TrainLoans, family = 'binomial')
step(Fmodel,scope=list(lower=Fmodel,upper=LrgModel),direction="both", trace=0)
stepAIC(Fmodel, direction = "both")
```

```{r echo=FALSE, warning=FALSE}

FLgrM =glm(formula = status ~ amount + term + rate + payment + grade + 
    length + debtIncRat + delinq2yr + inq6mth + openAcc + revolRatio + 
    totalAcc + totalPaid + totalBal + totalRevLim + bcOpen + 
    bcRatio + totalLim + totalRevBal + totalBcLim + totalIlLim, 
    family = "binomial", data = TrainLoans)

summary(FLgrM)
vif(FLgrM)
prLrg = predict(FLgrM, newdata = TestLoans, type = "response") # to do prediction
trs = 0.5#threshold of 0.5
prediction = cut(prLrg, breaks=c(-Inf,trs, Inf), labels=c("Bad", "Good"), header = TRUE)
PrEdCAL= table(prediction,TestLoans$status)
addmargins(PrEdCAL)

PLevel = round(sum(diag(PrEdCAL)) / sum(PrEdCAL)*100 ,2) # Percentage correctly predicted
bad = round(diag(PrEdCAL)[1]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate correctly predicted bad loans
good = round((PrEdCAL)[4]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate correctly predicted good loans
incbad = round((PrEdCAL)[3]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate incorrectly predicted bad loans
incgood = round((PrEdCAL)[2]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate incorrectly predicted good loans
print(paste0('Correctly predicted loans = ', PLevel,'%'))
print(paste0('Correctly predicted as Bad loans ', bad ,'% ', 'and as Good loans is ',good,'%'))
print(paste0('Incorrectly predicted as bad loans ', incbad,'% ', 'and as good loans is ',incgood,'%'))
```

# 9.Optimizing The Model  Threshold for Profit.

## 9.1. Threshold at 0.8 

```{r echo=FALSE, warning=FALSE}
prLrg = predict(FLgrM, newdata = TestLoans, type = "response") # to do prediction
trs = 0.8#threshold of 0.8
prediction = cut(prLrg, breaks=c(-Inf,trs, Inf), labels=c("Bad", "Good"), header = TRUE)
PrEdCAL= table(prediction,TestLoans$status)
addmargins(PrEdCAL)

PLevel = round(sum(diag(PrEdCAL)) / sum(PrEdCAL)*100 ,2) # Percentage correctly predicted
bad = round(diag(PrEdCAL)[1]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate correctly predicted bad loans
good = round((PrEdCAL)[4]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate correctly predicted good loans
incbad = round((PrEdCAL)[3]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate incorrectly predicted bad loans
incgood = round((PrEdCAL)[2]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate incorrectly predicted good loans
print(paste0('Correctly predicted loans = ', PLevel,'%'))
print(paste0('Correctly predicted as Bad loans ', bad ,'% ', 'and as Good loans is ',good,'%'))
print(paste0('Incorrectly predicted as bad loans ', incbad,'% ', 'and as good loans is ',incgood,'%'))
prediction = as.data.frame(prediction)
TestLoans = cbind(TestLoans,prediction)
Profit = TestLoans$totalPaid - TestLoans$amount 
TestLoans$Profit = Profit
profitG = TestLoans$Profit[TestLoans$prediction =='Good']
SUMP= round(sum(profitG))
SUMN=round(sum(Profit))
print(paste0("The normal profit is $",SUMN," and the predicted profit is $",SUMP))


```

##  9.2. Threshold at 0.3

```{r echo=FALSE, warning=FALSE}
prLrg = predict(FLgrM, newdata = TestLoans, type = "response") # to do prediction
trs = 0.3#threshold of 0.3
prediction2 = cut(prLrg, breaks=c(-Inf,trs, Inf), labels=c("Bad", "Good"), header = TRUE)
PrEdCAL= table(prediction2,TestLoans$status)
addmargins(PrEdCAL)

PLevel = round(sum(diag(PrEdCAL)) / sum(PrEdCAL)*100 ,2) # Percentage correctly predicted
bad = round(diag(PrEdCAL)[1]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate correctly predicted bad loans
good = round((PrEdCAL)[4]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate correctly predicted good loans
incbad = round((PrEdCAL)[3]*100/(diag(PrEdCAL)[1]+(PrEdCAL)[3]),2) # to calculate incorrectly predicted bad loans
incgood = round((PrEdCAL)[2]*100/((PrEdCAL)[2]+(PrEdCAL)[4]),2) # to calculate incorrectly predicted good loans
print(paste0('Correctly predicted loans = ', PLevel,'%'))
print(paste0('Correctly predicted as Bad loans ', bad ,'% ', 'and as Good loans is ',good,'%'))
print(paste0('Incorrectly predicted as bad loans ', incbad,'% ', 'and as good loans is ',incgood,'%'))
prediction2 = as.data.frame(prediction2)
TestLoans = cbind(TestLoans,prediction2)
Profit = TestLoans$totalPaid - TestLoans$amount 
TestLoans$Profit = Profit
profitG = TestLoans$Profit[TestLoans$prediction2 =='Good']
SUMP2= round(sum(profitG))
SUMN2=round(sum(Profit))
print(paste0("The normal profit is $",SUMN2," and the predicted profit is $",SUMP2))
```
